The development of children's ability to track and predict turn structure in conversation

### 2016-06-21 paper/Letters/ResponseLetter-Jun2016.tex

\noindent We thank the editor and our two reviewers for their very helpful comments on the most recently submitted manuscript. We have tried to address their comments here and in the manuscript.

\medskip

\noindent \textit{1. Are the permutation analyses anti-conservative?}

\noindent R2 found our new description of the permutation analysis satisfactory, but R1 still expressed some doubts that we believe may have arisen from a miscommunication. We therefore focused the description of the permutation analyses more on the purpose of the analyses (PXX). In particular, R1 expressed concern that the permutation analyses are anti-conservative because we assume that ``baseline saccades are equiprobable across the turn.'' While we appreciate R1's point, this characterization is not quite correct and we believe that our permutation analyses are doing exactly what they should: namely, testing the likelihood that we could get significant effects (of any predictor) if the saccades turned out to be randomly distributed---a true random baseline. Crucially, our permutations assume that baseline saccades are equiprobable across the entire stimulus, and not just across turns. R1 may imagine that we are comparing (a) anticipation rates at transition windows against (b) rates during turns, thereby maximizing our chances of finding a difference from the baseline. This is not what we are doing and is specifically what we criticize about the Keitel et al. (2013, 2015) studies. Instead, we compare (a) anticipation rates at transition windows to (b) transition rates over the whole stimulus. In other words, our baseline is agnostic to the placement of the true transition windows.

\noindent One possible response would be that presumably, most saccades are at transition windows and so, by comparing them to saccades over the whole stimulus, we are merely comparing them to at a watered down saccadic rate. We have two responses to this objection, however: First, the primary aim of the permutation analysis is actually to test R1's assumption (i.e., anticipation at transition windows is greater than elsewhere), but has the substantial added benefit of testing whether the assumption holds for each predictor effect in the mixed effect models. Second, the non-``transition window'' parts of the stimulus contain many behaviors that likely elicit anticipatory gaze behavior. For example, the transition windows in our real-turn analyses don't include: very early and very delayed anticipatory looks, looks to a responder in error (before the participant realizes the same speaker has actually continued speaking), and turn transitions that weren't included in our final analyses (e.g., because they had overlapped speech or very long inter-turn gaps). All three of these cases are common in our data and elicit similar looking behavior to the target ``transition window'' anticipatory gazes. But, because they are not part of our final ``real transition window'' dataset, they only count against us in the analyses---they \textit{only} appear in the baseline estimates, and never in the ``real'' transition estimates.

\noindent We therefore conclude that our permutation analyses are (a)  fit to the question we pose, (b) present a more useful and conservative baseline than used in prior literature, and (c) are conservative in the sense that there are other turn-relevant behaviors hidden in the non-``transition window'' parts of the stimulus. By editing the motivation behind the analyses on PXX, we hope we have made this argument clearer. R1 may still be dissatisfied with the permutation analyses in that they do not capture anything more than ``random-looker'' vs. ``turn-driven looker'' (in his/her words, that they are ``superfluous''), but this is exactly what they are intended to do. We do note, however, that we also test a more subtle hypothesis about anticipatory looking rates (i.e., ``boredom-driven looker'' vs. ``turn-driven looker'') in a separate set of analyses (Boredom Hypothesis; Appendix C, also discussed below).

@@

\noindent \textit{3. Are the permutation-based t-tests of age effects trustworthy?}

\noindent To supplement these analyses, we have added standard (non-permutation based) comparisons between the Normal speech and No speech (``party'') conditions for each age and have integrated them into the results for Expt 2.

@@

\noindent \textit{4. Shouldn't the effect of gap duration be part of the developmental story?}

\noindent R1 points out that, we should treat gap duration as more than just a control variable; that its effect is both strong and theoretically interesting. Following R1's helpful suggestion, we have added discussion about the role of inter-turn gaps for retroactive turn boundary assignment throughout the manuscript, from the Introduction to the General Discussion. We have worked to make it clear why gaps themselves can be a good cue to turn transition. We have also taken up R1's suggestion for reframing in the General Discussion; we now discuss the developmental trajectory with respect to both the condition-based linguistic effects and inter-turn silence, which more naturally integrates the effect of gap duration into our story (PXX). By doing this, we move ``away from the idea that linguistic prediction is the only explanatory factor'' as R1 suggests. R1 also noticed that we didn't include an age-by-duration or condition-by-duration interaction in our models. This is because of we were unable to add full interactions between gap duration and the other fixed effects in all models; in many models, higher-order interactions with gap duration create non-convergence issues due to data sparsity with some combinations of three or more predictor levels.

\noindent To address this issue, in our revision, we attempt to test the effect of gap duration but also remain consistent across all of our models. We added a fixed effect of gap duration and two-way interactions between gap duration and all other simple fixed effects (i.e., gap duration and age, linguistic condition, and transition type) to every model. This way all of our models test both the overall effect of gap duration (as before) and any interaction between gap duration and age, linguistic condition, or transition type. This tests the hypothesis that the gap duration effect might change with age (which, incidentally, we do not see in our data) as well as other potential effects of gap duration.

@@

\noindent \textit{5. Is there strong evidence that children are using linguistic information?}

\noindent R1 mentions that the linguistic effects in our study are relatively small, tucked within interactions. We take R1's point, but also expected from the start that linguistic effects would strongly interact with age (the reason we used a broad age range) and that anticipatory gaze would interact with speech act (the reason we included question coding). We did not, however, anticipate that the effect of speech act would dominate our results. The strong effects of speech act effectively funnels all the linguistic effects (and their age interactions) into the ``question'' subset of the data. Because the speech act effect is both new and theoretically motivated, and because there are still effects of age and linguistic condition associated with it, we believe that these findings are informative: they tell us about the conditions under which children and adults engage in spontaneous linguistic prediction about turn taking in conversation.

### 2016-06-23 paper/June2016-Submission/CasillasFrank-Anticip.tex
resubmitted manuscript

\begin{abstract}
Children begin developing turn-taking skills in infancy but take several years to assimilate their growing knowledge of language into their turn-taking behavior. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to turn structure. In Experiment 1, we showed English and non-English conversations to English-speaking adults and children. In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech. Children spontaneously made anticipatory gaze switches by age two and continued improving through age six. In both experiments, children and adults made more anticipatory switches after hearing questions. Consistent with prior findings on adult turn prediction, prosodic information alone did not increase children's anticipatory gaze shifts. But, unlike prior work with adults, lexical information alone was not sufficient either---children's performance was best overall with lexicosyntax and prosody together. Our findings support an account in which turn tracking and prediction emerges in infancy, and then only gradually becomes integrated with linguistic processing.
\end{abstract}

@@

Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different timing from adult turn taking, and lacks semantic content \citep{hilbrink2015, jaffe2001}. But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking. These early non-verbal interactions likely establish the protocol by which children come to use language with others. How then do children integrate linguistic knowledge with these preverbal turn-taking abilities?

In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how they integrate language into their predictions as they grow older. We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction. We then discuss recent work on the development of turn-taking skills before presenting the details of the present study.

@@

Adults have a lot of information at their disposal to help make accurate predictions. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreira2015}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. Additionally, the sequential context of a turn can make the next action obvious: answers after questions, thanks or denial after compliments, etc. \citep{schegloff2007}.

Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn. The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations. With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal. But when no lexical information was available, participants' responded significantly earlier within the turn. The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient. Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}). They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.

Despite this body of evidence, the role of prosody for adult turn prediction is still a matter of debate. De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody. Prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}. \citet*{torreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction. Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax. Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end. Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary. And when intonational phrase boundaries were embedded within multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time. These findings suggest that listeners actually do rely on prosodic cues to execute a response in interaction with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525). These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

@@

Before children manage to integrate linguistic cues into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions. For example, with little to no linguistic knowledge, children are often able to infer desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}. With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}. Thus, before they can anticipate turn structure form linguistic cues, children might react to silence as a cue to upcoming speaker change. Interactional silence itself may then serve as one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.

As children's language competence increases, they can use linguistic cues to make predictions about upcoming turn structure. Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions. Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}; they show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs. trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}. Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months infants can use prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}. In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.

@@

% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{torreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.

Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure. They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change. They showed their participants two types of videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech. Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and only 36-month-olds were affected by flattened intonation contours. This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically. They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension abilities (also see \citealp{keitel2015}).

@@

Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual. The average gap between turns was 900 msec, a duration much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}. The speakers in the videos were also asked to minimize their movements while performing scripted, adult-directed conversation, which would have created a somewhat unnatural interaction. Additionally, to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it difficult to distinguish the two talkers. Again, we attempt to address these issues in our current study. Despite these minor methodological drawbacks, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's ability to predict upcoming turn structure. Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}

@@

Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues across development. We present two experiments in which we measured children's anticipatory gaze to responders while they watched conversation videos with natural (people speaking English vs. non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues. We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment. We additionally tested for the use of one non-verbal cue: inter-turn silence.

We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a ``speech act'' effect). This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation. Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six, along with children's use of linguistic cues to anticipate answers after question turns. Third, children and adults often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood. Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding children's predictions (as is proposed for adults; \citealp{de-ruiter2006}). Anticipation is strongest for stimuli with the full range of linguistic cues. Our findings support an account in which turn prediction emerges in infancy, but becomes fully integrated with linguistic processing only gradually.

@@

We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation interspersed with attention-getting filler videos. Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean). Participants saw two videos in English and one in every other language. The participants, all native English speakers, were only expected to understand the two videos in English. We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., non-English prosody, gaze, in-breaths, phrase final lengthening). Using this method, we analyzed children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.

@@

We recruited 74 children ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment. We recruited adult participants through the Stanford University Psychology participant database. Adult participants were either paid or received course credit for their time. Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school and all of whom volunteered their time. All participants were native English speakers. Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with linguistic cues to turn taking. However, we are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample. The same applies to Experiment 2.} None of the adult participants reported fluency in a second language.

@@

After recording, we combined the audio and video recordings by hand, and cropped each one to the (approximate) 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language condition (English vs. non-English), inter-turn gap duration (in milliseconds), and transition type (question vs. non-question). Each non-English turn was coded as a question or non-question from a monolingual English-speaker's perspective, i.e., turns that ``sound like'' questions and turns that do not. We asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question. We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''. Thus, ``question'' cues in the non-English condition only resembled native English question cues, and therefore were likely harder to identify than cues to questionhood in the English condition. However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access in a foreign language? If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.

@@

\subsubsection*{Procedure}
Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display. The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height, approximately 60 cm from the viewer. We placed stereo speakers on the table, to the left and right of the display.

@@

As mentioned, most anticipatory switches happen in the inter-turn gap, but we also allowed anticipatory gaze switches that occurred in the final syllables of the prior turn. Early switches are consistent with the distribution of responses in explicit turn-boundary prediction tasks. For example, in a button press task, adult participants anticipated turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli came even earlier \citep{de-ruiter2006}. We therefore allowed switches to occur as early as 200 msec before the end of the prior turn. Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech. Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; Figure \ref{fig:criterion}). The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap). We did not count switches made during the fixation window as anticipatory. We \textit{did} count switches made during the inter-turn gap. The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (300 msec $+$ the inter-turn gap $+$ 300 msec).

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
\end{center}
\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition.}
\label{fig:criterion}
\end{figure}

>>>>

\begin{abstract}
Children begin developing turn-taking skills in infancy but take several years to fluidly integrate their growing knowledge of language into their turn-taking behavior. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling linguistic cues to turn structure. In Experiment 1, we showed English and non-English conversations to English-speaking adults and children. In Experiment 2, we phonetically controlled lexicosyntactic and prosodic cues in English-only speech. Children spontaneously made anticipatory gaze switches by age two and continued improving through age six. In both experiments, children and adults made more anticipatory switches after hearing questions. Consistent with prior findings on adult turn prediction, prosodic information alone did not increase children's anticipatory gaze shifts. But, unlike prior work with adults, lexical information alone was not sufficient either---children's performance was best overall with lexicosyntax and prosody together. Our findings support an account in which turn tracking and turn prediction emerge in infancy and then gradually become integrated with children's online linguistic processing.
\end{abstract}

@@

Children participate in sequential coordination (proto-turn taking) with their caregivers starting at three months of age---before they can rely on any linguistic cues \citep[see, among others, ][]{bateson1975, hilbrink2015, jaffe2001, snow1977}. However, infant turn taking is different from adult turn taking in several ways: it is heavily scaffolded by caregivers, has different inter-turn timing, and lacks semantic content \citep{hilbrink2015, jaffe2001}. But children's early, turn-structured social interactions are presumably a critical precursor to their later conversational turn taking, establishing the protocol by which children come to use language with others. How then do children integrate linguistic knowledge with these preverbal turn-taking abilities?

In this study, we investigate when children begin to make predictions about upcoming turn structure in conversation and how online linguistic processing becomes integrated into their predictions as they grow older. We first give a basic review of turn-taking research and the state of current knowledge about adult turn prediction. We then discuss recent work on the development of turn-taking skills before presenting the details of the present study.

@@

Adults have a lot of information at their disposal to help make accurate predictions. Lexical, syntactic, and prosodic information (e.g., \textit{wh}-words, subject-auxiliary inversion, and list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, bogelstorreira2015}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries and can sometimes act as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. Additionally, the sequential context of a turn can make the next action obvious: answers after questions, thanks or denial after compliments, etc. \citep{schegloff2007}.

Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure. De Ruiter and colleagues \citeyearpar{de-ruiter2006} asked participants to listen to snippets of spontaneous conversation and to press a button whenever they anticipated that the current speaker was about to finish his or her turn. The speech snippets were controlled for the amount of linguistic information present; some were normal, but others had flattened pitch, low-pass filtered speech, or further manipulations. With pitch-flattened speech, the timing of participants' button responses was comparable to their timing with the full linguistic signal. But when no lexical information was available, participants responded significantly earlier within the turn. The authors concluded that lexicosyntactic information\footnote{The ``lexicosyntactic'' condition only included flattened pitch and so was not exclusively lexicosyntactic---the speech would still have residual prosodic structure, including syllable duration and intensity.} was necessary and possibly sufficient for turn-end projection, while intonation was neither necessary nor sufficient. Congruent evidence comes from studies varying the predictability of lexicosyntactic and pragmatic content: adults anticipate turn ends better when they can more accurately predict the exact words that will come next (\citealp{magyari2012}; see also \citealp{magyari2014}). They can also identify speech acts within the first word of an utterance \citep{gisladottir2015}, allowing them to start planning their response at the first moment possible \citep{bogels2015}.

Despite this body of evidence, the role of prosody for adult turn prediction is still a matter of debate. De Ruiter and colleagues' (2006) experiment focused on the role of intonation, which is only a partial index of prosody. Prosody is tied closely to the syntax of an utterance, so the two linguistic signals are difficult to control independently \citep{ford1996}. \citet*{bogelstorreira2015} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-end prediction. Critically, their stimuli were cross-spliced so that each item had full prosodic cues to accompany the lexicosyntax. Because of the splicing, they were able to create items that had syntactically-complete units with no intonational phrase boundary at the end. Participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete phrase without an intonational phrase boundary. And when intonational phrase boundaries were embedded within multi-utterance turns, participants were tricked into pressing the ``turn-end'' button 29\% of the time. These findings suggest that listeners actually do rely on prosodic cues to execute a response, and that their use of prosodic cues interacts with their predictions about the unfolding syntactic structure (see also de \citet{de-ruiter2006}:525). These experimental findings corroborate other corpus and experimental work promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

@@

Before children manage to fully integrate linguistic processing into their turn-taking behaviors (for both turn prediction and production), they can rely on non-verbal interactional cues, including silence, eye gaze, body orientation, and gesture, to identify the boundaries of social actions. For example, with little to no linguistic knowledge, children are often able to infer desired responses to offers and requests by taking account of their interlocutor's non-verbal communicative behavior, the structure of routine events, and the affordances of the current interactional context \citep{reddy13, nomikou11, shatz78}. With respect to turn taking in particular, children's spontaneous vocalizations during interaction demonstrate a sensitivity to short inter-speaker gaps from infancy \citep{hilbrink2015}. Thus, before children can anticipate turn structure by integrating linguistic cues from unfolding speech, they might react to silence as a cue to upcoming speaker change. Interactional silence itself may then serve as one of children's first cues to turn structure, giving them information about when to respond before they can rely on language.

As children's language competence and speed of processing increases \citep{kail1991}, they become better equipped to use linguistic cues in making predictions about upcoming turn structure. Studies of early linguistic development point to a possible early advantage for prosody over lexicosyntax in children's turn-taking predictions. Infants can distinguish their native language's rhythm type from others soon after birth \citep{mehler1988, nazzi2003}. They also show preference for the typical stress patterns of their native language over others by 6--9 months (e.g., iambic vs. trochaic), and can use prosodic information to segment the speech stream into smaller chunks from 8 months onward \citep{johnson2001, morgan1995}. Four- to five-month-olds also prefer pauses in speech to be inserted at prosodic boundaries, and by 6 months infants can use prosodic markers to pick out sub-clausal syntactic units, both of which are useful for extracting turn structure from ongoing speech \citep{jusczyk1995, soderstrom2003}. In comparison, children show at best a very limited lexical inventory before their first birthday \citep{bergelson2013, shi2010}.

@@

% Prior work has focused mainly on lexicosyntax and intonation, and not on prosody proper (\citealp{de-ruiter2006}; \citealp{keitel2013}, but see \citealp{bogelstorreira2015}), even though infants seem to acquire the basic rhythmic properties of the prosodic signal first \citep{mehler1988, moon1993, nazzi2003}.

Keitel and colleagues \citeyearpar{keitel2013} were one of the first to explore how children use linguistic cues to predict upcoming turn structure. They asked 6-, 12-, 24-, and 36-month-old infants, and adult participants to watch short videos of conversation and tracked their eye movements at points of speaker change. They showed their participants two types of videos---one normal and one with flattened pitch---to test the role of intonation in participants' anticipatory predictions about upcoming speech. Comparing children's anticipatory gaze frequency to a random baseline, they found that only 36-month-olds and adults made anticipatory gaze switches more often than expected by chance, and that only 36-month-olds were affected by flattened intonation contours. This finding led Keitel and colleagues to conclude that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli lexicosemantically. They also suggest that intonation might play a secondary role in turn prediction, but only after children acquire more sophisticated, adult-like language comprehension skills (also see \citealp{keitel2015}).

@@

Second, the conversation stimuli \citet{keitel2013} used were somewhat unusual. The average gap between turns was 900 msec, a duration much longer than typical adult timing, which averages around 200 msec \citep{stivers2009}. The speakers in the videos were also asked to minimize their movements while performing scripted, adult-directed conversation, which would have created a somewhat unnatural interaction. Additionally, to produce more naturalistic conversation, it would have been ideal to localize the sound sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it difficult to distinguish the two talkers. Again, we attempt to address these issues in our current study. Despite these minor methodological drawbacks, the Keitel et al. \citeyearpar{keitel2013} study still demonstrates interesting age-based differences in children's predictions about upcoming turn structure. Our current work takes these findings as a starting point.\footnote{But also see \citet{casillas2012, casillas2013}.}

@@

Our goal in the current study is to find out when children begin to make predictions about upcoming turn structure and to understand how their predictions are affected by linguistic cues to turn taking across development. We present two experiments in which we measured children's anticipatory gaze to responders while they watched conversation videos with natural (people speaking English vs. non-English; Experiment 1) and non-natural (puppets with phonetically manipulated speech; Experiment 2) control over the presence of lexical and prosodic cues. We tested children across a wide range of ages (Experiment 1: 3--5 years; Experiment 2: 1--6 years), with adult control participants in each experiment. We additionally tested for the use of one non-verbal cue: inter-turn silence.

We highlight four primary findings: first, although children and adults use linguistic cues to make predictions about upcoming turn structure, they do so primarily to predict speaker transitions after questions (a ``speech act'' effect). This intriguing effect, which has not been reported previously, suggests that participants track unfolding speech for cues to upcoming speaker change, which may affect how they use linguistic cues more generally for anticipatory processing in conversation. Second, we find that children make more predictions than expected by chance starting at age two, but that this effect is small at first, and continues to improve through age six, along with children's use of linguistic cues to anticipate answers after question turns. Third, children and adults often used inter-turn silence (a non-verbal cue to turn structure) to make more predictive gaze switches to the responder, suggesting that non-verbal cues are useful for predicting turn structure early on and continue to be important in adulthood. Finally, we find no evidence for an early prosodic advantage in children's anticipations and, further, no evidence that lexical cues alone are comparable to the full linguistic signal in aiding children's predictions (as is proposed for adults; \citealp{de-ruiter2006}). Anticipation is strongest for stimuli with the full range of linguistic cues. Our findings support an account in which turn prediction emerges in infancy and becomes integrated with online linguistic processing gradually, possibly because of children's increased linguistic knowledge and speed of processing with development.

@@

We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation that were interspersed with attention-getting filler videos. Each conversation video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean). Participants saw two videos in English and one in every other language. The participants, all native English speakers, were only expected to understand the two videos in English. We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., non-English prosody, gaze, in-breaths, phrase final lengthening). Using this method, we analyzed children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.

@@

We recruited 74 children ages 3;0--5;11 and 11 undergraduate adults to participate in the experiment. We recruited adult participants through the Stanford University Psychology participant database. Adult participants were either paid or received course credit for their time. Our child sample included 19 three-year-olds, 32 four-year-olds, and 23 five-year-olds, all enrolled in a local nursery school and all of whom volunteered their time. All participants were native English speakers. Approximately one-third (N$=$25) of the children's parents and teachers reported that their child regularly heard a second (and sometimes third or further) language, but only one child frequently heard a language that was used in our non-English video stimuli, and we excluded his data from the analyses.\footnote{Multilingual children may make predictions about upcoming turn structure differently from their monolingual peers due to their more varied experiences with linguistic cues to turn taking. We are unable to test this hypothesis here due to the variability in multilingual language input and the diverse set of languages being learned in our sample. The same applies to Experiment 2.} None of the adult participants reported fluency in a second language.

@@

After recording, we combined the audio and video recordings by hand, and cropped each one to the (approximate) 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language condition (English vs. non-English), inter-turn gap duration (in milliseconds), and transition type (question vs. non-question). Each non-English turn was coded as a question or non-question from a monolingual English-speaker's perspective, i.e., turns that ``sound like'' questions and turns that do not. We asked five native American English speakers to listen to the audio recording for each non-English turn and judge whether it sounded like a question. We marked non-English turns as questions when at least 4 of the 5 listeners (80\%) said that the turn ``sounded like a question''. Thus, ``question'' cues in the non-English condition only \textit{resembled} native English question cues, and were therefore likely harder to identify than cues to questionhood in the English condition. However, since participants did not speak the non-English languages and would only ever treat ``question-sounding'' turns as questions, we proceeded with these analyses to see how pervasive question effects were---could they show up even without lexical access? If participants primarily rely on prosodic cues to question turns, it's possible that even non-English prosody can elicit anticipatory gaze switches for question-like turns.

@@

\subsubsection*{Procedure}
Participants sat in front of an SMI 120Hz corneal reflection eye-tracker mounted beneath a large flatscreen display. The display and eye-tracker were secured to a table with an ergonomic arm that allowed the experimenter to position the whole apparatus at a comfortable height and approximately 60 cm from the viewer. We placed stereo speakers on the table, to the left and right of the display.

@@

As mentioned, most anticipatory switches happen in the inter-turn gap, but we also allowed anticipatory gaze switches that occurred in the final syllables of the prior turn. Early switches are consistent with the distribution of responses in explicit turn-boundary prediction tasks. For example, in a button press task, adult participants anticipated turn ends approximately 200 msec in advance of the turn's end, and anticipatory responses to pitch-flattened stimuli came even earlier \citep{de-ruiter2006}. We therefore allowed switches to occur as early as 200 msec before the end of the prior turn. Again, because it takes 200 msec to plan an eye movement, we counted anticipatory switches, at the latest, 200 msec after the onset of speech. Therefore, for very early and very late switches, our requirement of 100 msec of fixation on each speaker would sometimes extend outside of the gaze launch window boundaries (200 msec before and after the inter-turn gap; dark gray boxes Figure \ref{fig:criterion}). The maximally available fixation window was therefore 100 msec before and after the earliest and latest possible switch point (300 msec before and after the inter-turn gap). We did not count switches made during the fixation window as anticipatory. We \textit{did} count switches made during the inter-turn gap. The period of time from the beginning of the possible fixation window on the prior speaker to the end of the possible fixation window on the responder was our total analysis window (300 msec $+$ the inter-turn gap $+$ 300 msec).

\begin{figure}[t]
\begin{center}
\includegraphics[width=0.7\textwidth]{figures/FIG-AnticipCriteria.png}
\end{center}
\caption{Schematic summary of the criteria for anticipatory gaze shifts from speaker A to speaker B during a turn transition. FIX = hypothetical fixation on speaker A or speaker B; dashed lines = hypothetical saccadic time.}
\label{fig:criterion}
\end{figure}
